---
title: "DATA 606 Data Project Proposal"
author: "Kai Lukowiak"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Data Preparation

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(data.table)
library(ggplot2)
library(ggthemes)
```

```{r  loading, cache=TRUE,  message=FALSE, warning=FALSE}
train <- fread("/Users/kailukowiak/Data606_Proposal/Zillow/train_2016.csv")
properties <- fread("/Users/kailukowiak/Data606_Proposal/Zillow/properties_2016.csv")
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

**Can baisic statistical models predict housing prices better than baisic markers like mean price for an area or inflation adjusted previous price?**


### Cases 

**What are the cases, and how many are there?**


### Data collection 

**Describe the method of data collection.**

Data collection was easy since the datasets were posted on kaggle.
Files can be dowloaded here: https://www.kaggle.com/c/zillow-prize-1/data


### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study based on previous housing prices and atributes. 

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

https://www.kaggle.com/c/zillow-prize-1/data

### Response 

**What is the response variable, and what type is it (numerical/categorical)?**

The response variable in numeric (log error or housing price error).

### Explanatory 

**What is the explanatory variable, and what type is it (numerical/categorival)?**

There are a mix of numeric and catigorical variables (nothing OLS can't handle)

### Relevant summary statistics 

**Provide summary statistics relevant to your research question. For example, if you’re comparing means across groups provide means, SDs, sample sizes of each group. This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r summarystats}
qplot(logerror, data =  train, bins = 400) + 
  ylab("Count") + 
  ggtitle("Counts of log error",subtitle = "For the Train Dataset") + 
  coord_cartesian(x = c(-0.5,0.5)) +
  theme_economist()
```


```{r}
glimpse(properties)
```

There are a lot of NAs.
```{r}
naVals <- properties %>% 
  summarise_all(funs(sum(is.na(.))/n())) %>% 
  gather(key = "Variable", value = "missingPercent")

```

```{r naChart, fig.width=6, fig.height=9}
naVals %>% 
  ggplot(aes(x = reorder( Variable, missingPercent), y = missingPercent)) + 
  geom_bar(stat = "identity") +
  ylim(0,1) +
  ggtitle("Percentage of Non- Missing Values") +
  coord_flip() +
  theme_economist()


```


